New insights into the Weddell Sea ecosystem applying a network approach

### 2022-11-16 Manuscript/Manuscript.Rmd
Abstract completed

---
title: New insights into the Weddell Sea ecosystem applying a network approach
journal: "`r rticles::copernicus_journal_abbreviations(journal_name = 'communication')`"
author:
  - given_name: Tomás I.
    surname: Marina
    affiliation: "1, *"
    email: tomasimarina@gmail.com
  - given_name: Leonardo A.
    surname: Saravia
    affiliation: "1, *"
    email: arysar@gmail.com
  - given_name: Susanne
    surname: Kortsch
    affiliation: 2
    # authors can have multiple affiliations, which can also be used to mark deceased coauthors
# If you have more than one corresponding author, add them manually using the following structure (note the commas):
# Two authors: Daniel Nüst (daniel.nuest@uni-muenster.de) and Josiah Carberry (j.carberry@orcid.org)
# Three authors or more: Daniel Nüst (daniel.nuest@uni-muenster.de), Josiah Carberry (j.carberry@orcid.org), and Markus Konkol (m.konkol@wwu.de)
# If the following line is uncommented, the "corresponding: true" above are ignored
correspongdingauthors: Tomás I. Marina (tomasimarina@gmail.com) and Leonardo A. Saravia (arysar@gmail.com)
# If authors contributed equally, please mark the respective author names with an asterisk '*' and add a further affiliation: 'These authors contributed equally to this work.'", see template.
affiliation:
  - code: 1
    address: Centro Austral de Investigaciones Científicas (CADIC-CONICET), Ushuaia, Argentina
  - code: 2
    address: University of Helsinki, Helsinki, Finland
  - code: "*"
    address: These authors contributed equally to this work.
abstract: |
  The abstract goes here.
  It can also be on _multiple lines_.
bibliography: ../WeddellSea.bib
running:
  title: New insights into the Weddell Sea food web
  author: Marina et al.
competinginterests: |
  The authors declare no competing interests.
# See https://publications.copernicus.org/for_authors/licence_and_copyright.html, normally used for transferring the copyright, if needed. 
# Note: additional copyright statements for affiliated software or data need to be placed in the data availability section. 
# copyrightstatement: |
### Note: unless stated otherwise, software and data affiliated with the manuscript are assumed to be published under the same licence as the article (currently Creative Commons 4.0)
availability:
  #  use this to add a statement when having only software code available
  #  use this to add a statement when having only data sets available
  #code: |
  #data: |
codedata: |
    R code and data set used in this work is available at https://github.com/EcoComplex/WeddellSea
#   sample: |
# videosupplement: |
authorcontribution: |
  TIM and LAS: Conceptualization (lead); Data curation (lead); Formal analysis (lead); Methodology (lead); Coding (lead); Writing – original draft (lead); Writing – review and editing (lead). SK: Conceptualization (lead); Formal analysis (supporting); Methodology (supporting); Coding (supporting); Writing – original draft (supporting); Writing – review and editing (supporting).
# disclaimer: |
  # We like Copernicus.
acknowledgements: |
  Thanks to the rticles contributors!
# appendix: |
#   \section{Figures and tables in appendices}
#   \subsection{Option 1}
#   If you sorted all figures and tables into the sections of the text, please also sort the appendix figures and appendix tables into the respective appendix sections.
#   They will be correctly named automatically.
#   \subsection{Option 2}
#   If you put all figures after the reference list, please insert appendix tables and figures after the normal tables and figures.
#   
#   `\appendixfigures` needs to be added in front of appendix figures
#   `\appendixtables` needs to be added in front of appendix tables
#   
#   Please add `\clearpage` between each table and/or figure. Further guidelines on figures and tables can be found below.
#   Regarding figures and tables in appendices, the following two options are possible depending on your general handling of figures and tables in the manuscript environment:
#   To rename them correctly to A1, A2, etc., please add the following commands in front of them:
output:
  rticles::copernicus_article: 
    highlight: NULL
    keep_tex: true
  bookdown::pdf_book:
    base_format: rticles::copernicus_article # for using bookdown features like \@ref()
editor_options: 
  markdown: 
    wrap: 72
---

>>>>

---
title: New insights into the Weddell Sea ecosystem applying a network approach
journal: "`r rticles::copernicus_journal_abbreviations(journal_name = 'communication')`"
author:
  - given_name: Tomás I.
    surname: Marina
    affiliation: "1, *"
    email: tomasimarina@gmail.com
  - given_name: Leonardo A.
    surname: Saravia
    affiliation: "1, *"
    email: arysar@gmail.com
  - given_name: Susanne
    surname: Kortsch
    affiliation: 2
    # authors can have multiple affiliations, which can also be used to mark deceased coauthors
# If you have more than one corresponding author, add them manually using the following structure (note the commas):
# Two authors: Daniel Nüst (daniel.nuest@uni-muenster.de) and Josiah Carberry (j.carberry@orcid.org)
# Three authors or more: Daniel Nüst (daniel.nuest@uni-muenster.de), Josiah Carberry (j.carberry@orcid.org), and Markus Konkol (m.konkol@wwu.de)
# If the following line is uncommented, the "corresponding: true" above are ignored
correspongdingauthors: Tomás I. Marina (tomasimarina@gmail.com) and Leonardo A. Saravia (arysar@gmail.com)
# If authors contributed equally, please mark the respective author names with an asterisk '*' and add a further affiliation: 'These authors contributed equally to this work.'", see template.
affiliation:
  - code: 1
    address: Centro Austral de Investigaciones Científicas (CADIC-CONICET), Ushuaia, Argentina
  - code: 2
    address: University of Helsinki, Helsinki, Finland
  - code: "*"
    address: These authors contributed equally to this work.
abstract: |
  The network approach has proved to shade light on the structure and stability of complex marine ecosystems. In recent years, such approach was succesfully applied to the study of polar ecosystems, improving our knowledge on how they might respond to ongoing environmental changes. The Weddell Sea, particularly the high continental shelf area (74-78ºS), is one of the most studied areas outside the Antarctic Peninsula in the Southern Ocean. However, few researches have considered the known complexity of the Weddell Sea. Understanding and projecting impacts of change in this ecosystem requires a framework considering the full complexity of it. Here we focus on the trophic interactions that underpin ecosystem structure and stability. We analysed the Weddell Sea food web, composed by 490 species and 16041 predator-prey interactions, which is the highest-resolved Antarctic food web constructed so far. The objective of our work was threefold: 1) estimate the strength for each interaction in the food web, 2) characterise the species considering weighted and unweighted properties, and 3) analyse the species' role in the stability of the food web. On the one hand, we found that the distribution of the interaction strength at the food web level is asymmetric, where weak interactions are prevalent. Including such information as a (weighted) property for species we detected a positive relationship between species mean interaction strength and two unweighted properties, trophic level and the total number of interactions. We also found that only few species are key in terms of food web stability, presenting particular weighted and unweighted properties. In the same analysis we have integrated food web and species information, enabling a more complete assessment of the ecosystem structure and function, likely highlighting the ecological processes at play in the Weddell Sea. We consider that our results provide new insights important for the development of effective policies and management strategies, particularly given the ongoing initiative to implement a Marine Protected Area (MPA) in the Weddell Sea.

bibliography: ../WeddellSea.bib
running:
  title: New insights into the Weddell Sea food web
  author: Marina et al.
competinginterests: |
  The authors declare no competing interests.
# See https://publications.copernicus.org/for_authors/licence_and_copyright.html, normally used for transferring the copyright, if needed. 
# Note: additional copyright statements for affiliated software or data need to be placed in the data availability section. 
# copyrightstatement: |
### Note: unless stated otherwise, software and data affiliated with the manuscript are assumed to be published under the same licence as the article (currently Creative Commons 4.0)
availability:
  #  use this to add a statement when having only software code available
  #  use this to add a statement when having only data sets available
  #code: |
  #data: |
codedata: |
    R code and data set used in this work is available at https://github.com/EcoComplex/WeddellSea
#   sample: |
# videosupplement: |
authorcontribution: |
  TIM and LAS: Conceptualization (lead); Data curation (lead); Formal analysis (lead); Methodology (lead); Coding (lead); Writing – original draft (lead); Writing – review and editing (lead). SK: Conceptualization (lead); Formal analysis (supporting); Methodology (supporting); Coding (supporting); Writing – original draft (supporting); Writing – review and editing (supporting).
# disclaimer: |
  # We like Copernicus.
acknowledgements: |
  Thanks to the rticles contributors!
# appendix: |
#   \section{Figures and tables in appendices}
#   \subsection{Option 1}
#   If you sorted all figures and tables into the sections of the text, please also sort the appendix figures and appendix tables into the respective appendix sections.
#   They will be correctly named automatically.
#   \subsection{Option 2}
#   If you put all figures after the reference list, please insert appendix tables and figures after the normal tables and figures.
#   
#   `\appendixfigures` needs to be added in front of appendix figures
#   `\appendixtables` needs to be added in front of appendix tables
#   
#   Please add `\clearpage` between each table and/or figure. Further guidelines on figures and tables can be found below.
#   Regarding figures and tables in appendices, the following two options are possible depending on your general handling of figures and tables in the manuscript environment:
#   To rename them correctly to A1, A2, etc., please add the following commands in front of them:
output:
  rticles::copernicus_article: 
    highlight: NULL
    keep_tex: true
  bookdown::pdf_book:
    base_format: rticles::copernicus_article # for using bookdown features like \@ref()
editor_options: 
  markdown: 
    wrap: 72
---

### 2022-11-16 Manuscript/Manuscript.tex
Abstract completed

\begin{abstract}
The abstract goes here. It can also be on \emph{multiple lines}.
\end{abstract}

@@

Finally, we run extinction simulations and estimated its impact on the
stability of the food web. For this, we calculated a stability index
called Quasi-Sign Stability (QSS), which is the proportion of stable
networks using randomized Jacobians and keeping the predator-prey sign
structure fixed \citep{Allesina2008}. If this proportion is zero, then
one should take into account the real part of the maximum eigenvalue of
the Jacobian matrix, which is also a measure of stability
\citep{Grilli2016}.

@@

It's important to note that since the proportion of Jacobians that are
locally stable or QSS was zero for the Weddell Sea food web, we
considered the mean maximum eigenvalue as the stability index. When
performing the extinction simulations, our results showed that the
majority of the species had no significant impact on the stability of
the food web when extinct (Figure 5). This is shown in Figure 5, where
most of the points lie around the zero value of the `Stability
difference' y axis, meaning that the stability for the food web minus a
given species is similar (not significant) to the stability for the
whole network. However, there were few species that when become extinct
altered significantly the stability of the food web. Most of such
species generated positive significant differences in the stability of
the food web, which mean that they increased the network's stability
when deleted. There were two exceptions, Pagetopsis macropterus
(demersal fish) and Maxilliphimedia longipes (benthopelagic amphipod),
that when deleted decreased the stability. Overall, 15 out of 490
species (3.06\%) gave rise to significant changes in the food web's
stability (Table 2).

>>>>

\begin{abstract}
The network approach has proved to shade light on the structure and
stability of complex marine ecosystems. In recent years, such approach
was succesfully applied to the study of polar ecosystems, improving our
knowledge on how they might respond to ongoing environmental changes.
The Weddell Sea, particularly the high continental shelf area (74-78ºS),
is one of the most studied areas outside the Antarctic Peninsula in the
Southern Ocean. However, few researches have considered the known
complexity of the Weddell Sea. Understanding and projecting impacts of
change in this ecosystem requires a framework considering the full
complexity of it. Here we focus on the trophic interactions that
underpin ecosystem structure and stability. We analysed the Weddell Sea
food web, composed by 490 species and 16041 predator-prey interactions,
which is the highest-resolved Antarctic food web constructed so far. The
objective of our work was threefold: 1) estimate the strength for each
interaction in the food web, 2) characterise the species considering
weighted and unweighted properties, and 3) analyse the species' role in
the stability of the food web. On the one hand, we found that the
distribution of the interaction strength at the food web level is
asymmetric, where weak interactions are prevalent. Including such
information as a (weighted) property for species we detected a positive
relationship between species mean interaction strength and two
unweighted properties, trophic level and the total number of
interactions. We also found that only few species are key in terms of
food web stability, presenting particular weighted and unweighted
properties. In the same analysis we have integrated food web and species
information, enabling a more complete assessment of the ecosystem
structure and function, likely highlighting the ecological processes at
play in the Weddell Sea. We consider that our results provide new
insights important for the development of effective policies and
management strategies, particularly given the ongoing initiative to
implement a Marine Protected Area (MPA) in the Weddell Sea.
\end{abstract}

@@

Finally, we run extinction simulations and estimated its impact on the
stability of the food web. For this, we calculated a stability index
called Quasi-Sign Stability (QSS), which is the proportion of stable
networks using randomized Jacobians and keeping the predator-prey sign
structure fixed \citep{Allesina2008}. If this proportion is zero, then
one should take into account the real part of the maximum eigenvalue of
the Jacobian matrix, which is also a measure of stability
\citep{Grilli2016}, indicating a more stable food web when it is
negative.

@@

It's important to note that since the proportion of Jacobians that were
locally stable or QSS was zero for the Weddell Sea food web, we
considered the mean maximum eigenvalue as the stability index. When
performing the extinction simulations, our results showed that the
majority of the species had no significant impact on the stability of
the food web when extinct (Figure 5). This is shown in Figure 5, where
most of the points lie around the zero value of the `Stability
difference' y axis, meaning that the stability for the food web minus a
given species is similar (not significant) to the stability for the
whole network. However, there were few species that when become extinct
altered significantly the stability of the food web. Most of such
species generated positive significant differences in the stability of
the food web, which mean that they increased the network's stability
when deleted. There were two exceptions, Pagetopsis macropterus
(demersal fish) and Maxilliphimedia longipes (benthopelagic amphipod),
that when deleted decreased the stability. Overall, 15 out of 490
species (3.06\%) gave rise to significant changes in the food web's
stability (Table 2).
